---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```
[![Travis-CI Build Status](https://travis-ci.org/hrbrmstr/libstableR.svg?branch=master)](https://travis-ci.org/hrbrmstr/libstableR)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/hrbrmstr/libstableR?branch=master&svg=true)](https://ci.appveyor.com/project/hrbrmstr/libstableR)

# libstableR

Fast and Accurate Evaluation, Random Number Generation and Parameter Estimation of Skew Stable Distributions

`libstableR` provides functions to work with skew stable distributions in a fast and accurate way. It performs:

* Fast and accurate evaluation of the probability density function (PDF) and cumulative density function (CDF).
* Fast and accurate evaluation of the quantile function (CDF^{-1}).
* Random numbers generation.
* Skew stable parameter estimation with:
      * McCulloch's method of quantiles.
      * Koutrouvellis' method.
      * Maximum likelihood estimation.
      * Modified maximum likelihood estimation.
* The evaluation of the PDF and CDF is based on the formulas provided by John P Nolan.

## Authors

- Javier Royuela del Val <jroyval@@lpi.tel.uva.es> (Maintainer)
- Federico Simmross Wattenberg
- Carlos Alberola López

(I just rounded out some corners and added macOS compatiblity)

## What's in the tin?

- `stable_cdf`:	PDF and CDF of a skew stable distribution.
- `stable_fit`:	Methods for parameter estimation of skew stable distributions.
- `stable_fit_init`:	Methods for parameter estimation of skew stable distributions.
- `stable_fit_koutrouvelis`:	Methods for parameter estimation of skew stable distributions.
- `stable_fit_mle`:	Methods for parameter estimation of skew stable distributions.
- `stable_fit_mle2d`:	Methods for parameter estimation of skew stable distributions.
- `stable_pdf`:	PDF and CDF of a skew stable distribution.
- `stable_pdf_and_cdf`:	PDF and CDF of a skew stable distribution.
- `stable_q`:	Quantile function of skew stable distributions
- `stable_rnd`:	Skew stable distribution random sample generation.

## Usage

```{r message=FALSE, warning=FALSE}
# devtools::install_github("hrbrmstr/libstableR") # macOS compatibility
# install.packages("libstableR") # macOS compat not on CRAN yet

library(libstableR)

# Set alpha, beta, sigma and mu stable parameters in a vector
pars <- c(1.5, 0.9, 1, 0)

# Generate an abscissas axis and probabilities vector
x <- seq(-5, 10, 0.05)
p <- seq(0.01, 0.99, 0.01)

# Calculate pdf, cdf and quantiles
pdf <- stable_pdf(x, pars)
head(pdf)

cdf <- stable_cdf(x, pars)
head(cdf)

xq  <- stable_q(p, pars)
head(xq)

# Generate 300 random values
rnd <- stable_rnd(300, pars)
head(rnd)

# Estimate the parameters of the skew stable distribution given
# the generated sample:

# Using the McCulloch's estimator:
pars_est_M <- stable_fit_init(rnd)
head(pars_est_M)

# Using the Koutrouvelis' estimator:
pars_est_K <- stable_fit_koutrouvelis(rnd, pars_est_M)
head(pars_est_K)
```

```{r plot, message=FALSE, warning=FALSE, fig.retina=2}
alpha <- 1.25;  # stability index
beta  <- 0.95;  # skew index
sigma <- 2.0;   # scale parameter
mu    <- 0.0;   # location parameter

pars <- c(alpha, beta, sigma, mu)

# abscissae vector
x    <- seq(from=-5, to=100, by=0.1)

# evaluation of pdf and cdf
pdf  <- stable_pdf(x, pars, parametrization=0, tol=1e-12)
cdf  <- stable_cdf(x, pars, parametrization=0, tol=1e-12)

# probabilities vector and quantile function evaluation
p  <- seq(from=0.01, to=0.99, by=0.01)
q  <- stable_q(p, pars, parametrization=0, tol=1e-12)

# Random numbers generation
N   <- 300;
rnd <- stable_rnd(N, pars, seed=1234)

# Parameter estimation with Koutrouvelis method.
# McCulloch estimator will be used for initialization.
pars_init <- stable_fit_init(rnd)
pars_K    <- stable_fit_koutrouvelis(rnd, pars_init)

# Get the pdf given by the estimated parameters
pdf_init <- stable_pdf(x, pars_init)
pdf_K    <- stable_pdf(x, pars_K)

# Plot histogram, true pdf and pdf from the estimated parameters
xlim <- c(-5, 30)
ylim <- c(0, 0.15)
breaks <- seq(min(rnd)-1, max(rnd)+1, 0.5);
hist(rnd, breaks=breaks, freq=FALSE, xlim=xlim, ylim=ylim)

lines(x, pdf, col = "red", lwd=2)
lines(x, pdf_init, col = "green", lwd=2)
lines(x, pdf_K, col = "blue", lwd=2)

legend(20, 0.15,
       c("Histogram", "True pdf", "McCulloch", "Koutrouveils"),#,"Max. likelihood"),
       lty = c(1,1,1,1,1),
       lwd = c(2,2,2,2,2),
       col=c("black","red","green","blue"))#,"yellow"))
```
